Deep mutational scanning of hemagglutinin helps distinguish the evolutionary fate of human H3N2 influenza virus lineages

### 2018-01-16 paper/manuscript.tex
Re-edit Introduction

Two main features distinguish the evolutionarily successful trunk from its competitors: greater antigenic change, and efficient viral growth and transmission.
In principle, experiments could be informative for identifying how mutations affect these features.
Most work on influenza evolution to date has utilized experimental data primarily to assess the antigenicity of circulating strains~\cite{sun2013using,harvey2016identification,neher2016prediction}.
While the antigenic features of a virus certainly contribute to its success, the non-antigenic effects of mutations also play an important role~\cite{pybus2007phylogenetic,kucharski2011influenza,strelkowa2012clonal,luksza2014predictive,koelle2015effects}.
Specifically, due to influenza virus's high mutation rate~\cite{holland1982rapid,steinhauer1987rapid,lauring2010quasispecies} and lack of intra-segment recombination~\cite{boni2008homologous}, deleterious mutations become linked to beneficial ones.
The resulting accumulation of deleterious mutations can affect non-antigenic properties that are central to viral fitness~\cite{luksza2014predictive}.
However, while extensive work has gone into characterizing how mutations affect HA antigenicity~\comment{epitope site papers, Koel paper, Hensley papers, Kawaoka paper}, we lack any large-scale quantitative characterization of how mutations to H3N2 HA affect non-antigenic properties crucial to viral fitness.

@@

We performed a similar analysis using the current data for the Perth/2009 H3 HA.
Surprisingly, there was much less contrast in mutational tolerance between the stalk and head domains for the H3 HA than for the H1 (Figure~\ref{fig:mut_tolerance}).
For instance, in the H3 HA, the short helix A in the stalk is very mutationally tolerant.
Interestingly, more studies have reported selecting escape mutants from broadly neutralizing anti-stalk antibodies in H3~\cite{ekiert2011highly, friesen2014common, chai2016two, yamayoshi2017human} than in H1~\cite{okuno1993common,doud2017quantifying} HAs.

>>>>

Two main characteristics distinguish the evolutionarily successful trunk from its competitors: greater antigenic change, and efficient viral growth and transmission.
In principle, experiments could be informative for identifying how mutations affect these features.
Most work on influenza evolution to date has utilized experimental data primarily to assess the antigenicity of circulating strains~\cite{sun2013using,harvey2016identification,neher2016prediction}.
While the antigenic features of a virus certainly contribute to its success, the non-antigenic effects of mutations also play an important role~\cite{pybus2007phylogenetic,kucharski2011influenza,strelkowa2012clonal,luksza2014predictive,koelle2015effects}.
Specifically, due to influenza virus's high mutation rate~\cite{holland1982rapid,steinhauer1987rapid,lauring2010quasispecies} and lack of intra-segment recombination~\cite{boni2008homologous}, deleterious mutations become linked to beneficial ones.
The resulting accumulation of deleterious mutations can affect non-antigenic properties that are central to viral fitness~\cite{luksza2014predictive}.
However, while extensive work has gone into characterizing how mutations affect HA antigenicity~\cite{laver1979antigenic, webster1980determination,wiley1981structural,koel2013substitutions,chambers2015identification,li2016selection}, we lack any large-scale quantitative characterization of how mutations to H3N2 HA affect non-antigenic properties crucial to viral fitness.

@@

We performed a similar analysis using the current data for the Perth/2009 H3 HA.
Surprisingly, there was much less contrast in mutational tolerance between the stalk and head domains for the H3 HA than for the H1 (Figure~\ref{fig:mut_tolerance}).
For instance, in the H3 HA, the short helix A in the stalk is very mutationally tolerant.
Interestingly, more studies have reported selecting escape mutants from broadly neutralizing anti-stalk antibodies in H3~\cite{ekiert2011highly, friesen2014common, chai2016two, yamayoshi2017human} than in H1~\cite{okuno1993common,doud2017quantifying,anderson2017natural} HAs.
